---
title: "Project 1"
author: "Name:  \n Partner: "
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 3
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}
---
```{r setup, include = FALSE}
#### Load necessary packages ####
# * These packages are not necessary to complete the assignment and or only used 
#   to provide an example. 
packages <- c("knitr", "kableExtra", "magrittr", "readr", "geosphere")
install_me <- packages[!(packages %in% installed.packages()[, "Package"])]
if (length(install_me)) install.packages(install_me)
library(knitr)
library(kableExtra)
library(magrittr)
library(readr)
library(geosphere)
library(dplyr)
```

## Background
The World Health Organization has recently employed a new data science initiative, *CSIT-165*, that uses data science to characterize pandemic diseases. 
*CSIT-165* disseminates data driven analyses to global decision makers.

*CSIT-165* is a conglomerate comprised of two fabricated entities: *Global Health Union (GHU)* and *Private Diagnostic Laboratories (PDL)*. 
Your and your partner's role is to play a data scientist from one of these two entities.

### Morgan Henderson 
will be part of the GHU and will provide time series data counting COVID-19 related recoveries and and deaths.

### Preena Maruthavelu
will be part of the PDL and will provide time series data counting COVID-19 related confirmations.

## Data
> [2019 Novel Coronavirus COVID-19 (2019-nCoV) Data Repository by John Hopkins CSSE](https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series)
Data for 2019 Novel Coronavirus is operated by the John Hopkins University Center for Systems Science and Engineering (JHU CSSE).
Data includes daily time series CSV summary tables, including confirmations, recoveries, and deaths. 
Country/region are countries/regions hat conform to World Health Organization (WHO).
Lat and Long refer to coordinates references for the user. 
Date fields are stored in MM/DD/YYYY format.

## Project Objectives

### Objective 1

```{r ob1}

confirmed_cases <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"
covid_deaths <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"
#load data
cases_df <- read.csv(confirmed_cases, header = TRUE, na.strings = c("", " "))
deaths_df <- read.csv(covid_deaths, header = TRUE, na.strings = c("", " "))

#segment first day of COVID data
data_cases <- dplyr::select(cases_df, Province.State, Country.Region, X1.22.20); 
data_deaths <- dplyr::select(deaths_df, Province.State, Country.Region, X1.22.20)

# Filter for the first day and select relevant columns
first_day_cases <- cases_df %>%
  filter(X1.22.20 != 0) %>%
  select(Province.State, Country.Region, X1.22.20)

first_day_deaths <- deaths_df %>%
  filter(X1.22.20 != 0) %>%
  select(Province.State, Country.Region, X1.22.20)

# Identify the country with the highest confirmed cases and highest deaths
max_cases <- first_day_cases %>%
  filter(X1.22.20 == max(X1.22.20)) %>%
  pull(Country.Region)

max_deaths <- first_day_deaths %>%
  filter(X1.22.20 == max(X1.22.20)) %>%
  pull(Country.Region)
# Determine if the country is the origin of the outbreak
if(max_cases == max_deaths) {
  output <- paste("The origin of the COVID-19 outbreak was likely", max_cases)
  print(output)
}

```

### Objective 2
```{r ob2}
```

### Objective 3
```{r ob3}
```

### Objective 4

#### Objective 4.1
```{r ob4.1}
```

#### Objective 4.2
```{r ob4.2}
```

### GitHub Log
```{bash gitlog} 
git log --pretty=format:"%nSubject: %s%nAuthor: %aN%nDate: %aD%nBody: %b"
```
